Vagueness as Cost Reduction

### 2016-07-30 papers/2016/vcr.tex
got a working version of a 2016 draft in apa6 format

\documentclass[doc,apacite]{apa6}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{booktabs}
\usepackage{layout}
\usepackage{apacite}
\usepackage{datetime}
\usepackage{rotating}
\title{Vagueness as cost reduction}
\date{\today}
\authornote{\today ~\currenttime}
\shorttitle{Vagueness as cost reduction}
\twoauthors{Matt Green}{Kees van Deemter}
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}

\abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations. We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness. We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}

\begin{document}
\maketitle               
\section{Introduction}
Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a more particular meaning. Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \citeA{EgreKlinedinst}, among many others).  The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}. A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}. By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}. Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport. Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall''). Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs. The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative. These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.

@@

Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}. Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}. For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''. We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis. The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237''). This predicts that whenever a vague expression can perform the same communicative task as a precise one, it is rational to choose the vague expression. The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.

Questions concerning optimal language use have many practical applications. Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite<>[]{reiter2000building}. The process parallels language production in humans.}  (henceforth NLG) systems must make decisions between different formulations of the same information. For example, if a man's height is 6 foot 2 inches, this could be expressed as ``187.96 metres'', ``6 foot 2'', or ``tall'', among other ways, and the NLG system must decide between these.  The problem is particularly relevant for NLG systems that take numbers as input, as many do. In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc. \cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}. At present, such NLG systems are often forced to make decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition. A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.

@@

In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss. In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index). They were told that an ideal HHI score lies in the range of 45 to 55. In a longitudinal study, participants submitted their height, weight, hydration level, gender, and age to a computer each week. Participants were told that two algorithms would be used to compute their HHI, and that it was possible that the two algorithms might give different values initially, but would converge over the course of the study to a single value. They were also told that if the two algorithms did give different values, then the true score lay between the two values. In one condition, which the authors called the precise condition, the two algorithms gave the same score. In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.

One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values. However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores. The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores. This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance. In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}. These results are interesting, and of obvious potential practical importance. We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases. In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}. Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.

Similar issues arise from the work of \citeA{peters2009bringing}. The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care. There was a between-subjects manipulation based on numeracy. The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories). Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings. For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.

@@

\label{accann}
Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and gap size and their interaction as independent variables, and with random slopes for vagueness and gap size over participants. 
%
The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
%
Error rates decreased as gap size increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$). 
%
Error rates were greater in the crisp conditions than the vague conditions when gap size was small, and this difference diminished with increasing gap size until it reversed at the biggest gap size.
\begin{figure}[htbp]
\fitfigure{images/resultse1.pdf}
\caption{Experiment 1 results. }
\label{resultse1}
\end{figure}

@@

\begin{table}[htbp]
\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
\label{instructionse2}
\begin{tabular}{rlll}
\toprule
vagueness&instruction format&selection task&example\\
\midrule
crisp 	& 	numeric	& matching	&Choose the square with 6 dots \\
vague	&	numeric & matching	&Choose a square with about 10 dots\\
crisp	&	verbal	& comparison	&Choose the square with the fewest dots\\
vague	&	verbal	& comparison	&Choose a square with few dots\\
\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
\caption{Response time results for Experiment 2}
\label{resultse2}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
\caption{Distribution of Nearest match, borderline case, and furthest match responses}
\label{borderline-response-distribution-e2}
\end{figure}

@@

The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies. The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number. We used the same stimuli as in experiment 2. Table \ref{Instructions for e3} shows the instructions. % KvD I think we should motivate our choice of materials (especially "far ..er")

\begin{table}[htbp]
\centering
\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
\label{Instructions for e3}
\begin{tabular}{llll}
\hline
vagueness&instruction format&selection task&instruction\\
\hline
crisp & numeric&matching & Choose a square with 6 dots \\ 
crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
vague & numeric&matching & Choose a square with about 10 dots \\ 
vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
\hline
\end{tabular}
\end{table}%

@@

38 volunteers were recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds each for participating.
The apparatus was the same as Experiment 1.
The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
Each stimulus was an instruction followed by a triple of dots.
First a referring expression instruction was presented. Participants pressed a key to dismiss the instruction and proceed to the squares with dots in them.

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
\caption{Results for Experiment 3}
\label{resultse3}
\end{figure}

\section{Experiment 4}
\subsection{Introduction}
This experiment investigated response times for instructions that did not use a number. 
We manipulated vagueness and the selection task (comparison and matching). 
In order to implement the experiment without mentioning numbers,  we used a prime to visually show the numbers of dots that we wanted to refer to as {\em the target} in the instructions. 
This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \citeA{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.
An item was thus a combination of a visual prime, a numeric triple, and a referring expression.
The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e4}.

\begin{table*}[htbp]
\centering
\caption{Experiment 4: Instructions}
\label{Instructions for e4}
\begin{tabular}{lllp{7cm}}
\hline
vagueness&instruction format& selection task&instruction\\
\hline
crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
crisp & verbal&comparison& Choose a square with fewer dots than the target \\
vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
vague & verbal&compaison& Choose a square with far fewer dots than the target \\ 
\hline
\end{tabular}
\end{table*}%

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
\caption{Results for Experiment 4}
\label{resultse4}
\end{figure}

@@

\bibliography{vagueness-dec-2015}
\clearpage
\appendix

\begin{sidewaystable}
\centering
\caption{All Instructions}
\label{All Instructions}
\begin{tabular}{lp{5cm}llllp{2cm}l}
\midrule
Experiment&Example instruction&vagueness&quantifier type&symmetry&selection&number of valid targets&definiteness\\
\midrule
1&Choose the square with 5 dots&crisp&numeric&symmetric&matching&1&definite\\
1&Choose the square with few dots&vague&linguistic&asymmetric&comparison&1&definite\\
\midrule
2&Choose the square with 6 dots&crisp&numeric&symmetric&matching&1&definite\\
2&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
2&Choose the square with the fewest dots&crisp&linguistic&asymmetric&comparison&1&definite\\
2&Choose a square with few dots&vague&linguistic&asymmetric&comparison&2&indefinite\\
\midrule
3&Choose a square with 6 dots&crisp&numeric&symmetric&matching&1&indefinite\\
3&Choose a square with fewer than 20 dots&crisp&numeric&asymmetric&comparison&2&indefinite\\
3&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
3&Choose a square with far fewer than 20 dots&vague&numeric&asymmetric&comparison&1&indefinite\\
\midrule
4&Choose a square with the same number of dots as the target&crisp&linguistic&symmetric&matching&1&indefinite\\
4&Choose a square with fewer dots than the target&crisp&linguistic&asymmetric&comparison&2&indefinite\\
4&Choose a square with about the same number of dots as the target&vague&linguistic&symmetric&matching&2&indefinite\\
4&Choose a square with far fewer dots than the target&vague&linguistic&asymmetric&comparison&1&indefinite\\
\midrule
\end{tabular}
\end{sidewaystable}
\end{document}

>>>>

\documentclass[
a4paper 
%, man
, doc
%, jou
, draftfirst 
, longtable
]{apa6} 
\usepackage{rotating} 
                          % use this in jou and doc modes only, with \sideways. 
                          % In man mode use the documentclass option longtable and do not load rotating.
\usepackage{csquotes}  
\usepackage[american]{babel}                     
\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, doi=false, isbn=false]{biblatex}
    \DeclareLanguageMapping{american}{american-apa}
    \addbibresource{vcr.bib}
\usepackage{datetime}
\usepackage{url}

                         %% OTHER DOCUMENTCLASS OPTIONS 
                         % floatsintext (in man mode do not flush floats, better for internal reading, 
                         %                    but do not use for submission to journals)
                         % mask (masks references that are marked as self-references, for blind peer review)
                         % longtable (in man mode only)
                         % notxfonts (in jou mode suppresses txfonts in case pslatex or times is preferred)
                         % notimes (in jou mode use Computer Modern instead of loading txfonts or pslatex or times)
                         % notab (in jou mode suppress auto stretching of tabular environments to fill their float)
                         % helv (in man mode use Helvetica instead of Computer Modern)
                         % tt (in man mode use typewriter font)
                         % draftfirst or draftall (places the word DRAFT on first page or all pages)
                         % draft (is passed to, and handled by, the article documentclass)

\SetWatermarkAngle{0} % Angle at which the watermark text is drawn
\SetWatermarkColor{red} % Color of the watermark
%\SetWatermarkLightness{?real?} Lightness of the watermark text (1=white, 0=black)
%\SetWatermarkFontSize{?length?} Font size of the watermark text
\SetWatermarkScale{0.5} % Scaling of the watermark text 
%\SetWatermarkHorCenter{?length?} Horizontal center of watermark text 
\SetWatermarkVerCenter{2cm} % Vertical center of watermark text
\SetWatermarkText{Internal Draft} %Watermark text

\title{Vagueness as Cost Reduction}
\shorttitle{Vagueness as Cost Reduction}
\twoauthors{Matt Green}{Kees van Deemter}
\leftheader{Green, van Deemter} % for even-page headers in jou mode
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
\abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations. We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness. We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}
\keywords{keyword 1, keyword2, keyword3, keyword4}
\authornote{Contact author is Matt Green mjglab@gmail.com} % The Author Note, containing contact information, acknowledgements, etc.
\note{\date{\today~\currenttime}} % Notation of manuscript date or other information de- sired beneath affiliation line
\journal{Aimed at Language, Cognition, and Neuroscience} % typeset in the top left header of page 1 (jou and doc modes only)
\volume{volume info} % Volume, number, pages; typeset in the top left header in jou and doc modes, underneath the content of \journal
\ccoppy{Copyright notice etc} % Copyright notice, etc.; typeset in the top right header of page 1 (jou and doc modes only)
\copnum{Additional text} % Any additional text needed;  typeset in the top right header in jou and doc modes, underneath the content of \ccoppy

\begin{document}
\maketitle %
%\section{Introduction} % APA does not use Introduction header
Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons (see e.g. the entry ``vague'' in \cite{penguin}.) 
In most academic use though, the word `vagueness' has a more particular meaning. Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \textcite{EgreKlinedinst}, among many others).  The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}. A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}. By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite[][p.~1]{lipmanvague}. Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport. Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall''). Lipman uses this scenario to explain why standard game theory models of communication \cite[e.g.,]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs. The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative. These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.

@@

Several tentative answers to Lipman's question have been offered  \parencite[see][]{van2009utility, van2010vagueness}. Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}. For example, \textcite[][p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''. We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis. The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237''). This predicts that whenever a vague expression can perform the same communicative task as a precise one, it is rational to choose the vague expression. The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.

Questions concerning optimal language use have many practical applications. Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite[][]{reiter2000building}. The process parallels language production in humans.}  (henceforth NLG) systems must make decisions between different formulations of the same information. For example, if a man's height is 6 foot 2 inches, this could be expressed as ``187.96 metres'', ``6 foot 2'', or ``tall'', among other ways, and the NLG system must decide between these.  The problem is particularly relevant for NLG systems that take numbers as input, as many do. In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc. \cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}. At present, such NLG systems are often forced to make decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition. A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.

@@

In a series of studies of behaviour modification, \textcite[]{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss. In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index). They were told that an ideal HHI score lies in the range of 45 to 55. In a longitudinal study, participants submitted their height, weight, hydration level, gender, and age to a computer each week. Participants were told that two algorithms would be used to compute their HHI, and that it was possible that the two algorithms might give different values initially, but would converge over the course of the study to a single value. They were also told that if the two algorithms did give different values, then the true score lay between the two values. In one condition, which the authors called the precise condition, the two algorithms gave the same score. In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.

One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values. However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores. The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores. This ``illusion of proximity'' \cite[][p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance. In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}. These results are interesting, and of obvious potential practical importance. We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases. In the terminology of \textcite{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}. Furthermore, the experiments of \textcite{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.

Similar issues arise from the work of \textcite{peters2009bringing}. The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care. There was a between-subjects manipulation based on numeracy. The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories). Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings. For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.

@@

\label{accann}
Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and gap size and their interaction as independent variables, and with random slopes for vagueness and gap size over participants. 
%
The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
%
Error rates decreased as gap size increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$). 
%
Error rates were greater in the crisp conditions than the vague conditions when gap size was small, and this difference diminished with increasing gap size until it reversed at the biggest gap size.
\begin{figure*}[htbp]
\fitfigure{images/resultse1.pdf}
\caption{Experiment 1 results. }
\label{resultse1}
\end{figure*}

@@

\begin{table*}[htbp]
\begin{center}
\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
\label{instructionse2}
\begin{tabular}{rlll}
\toprule
vagueness&instruction format&selection task&example\\
\midrule
crisp 	& 	numeric	& matching	&Choose the square with 6 dots \\
vague	&	numeric & matching	&Choose a square with about 10 dots\\
crisp	&	verbal	& comparison	&Choose the square with the fewest dots\\
vague	&	verbal	& comparison	&Choose a square with few dots\\
\bottomrule
\end{tabular}
\end{center}
\end{table*}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
\caption{Response time results for Experiment 2}
\label{resultse2}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
\caption{Distribution of Nearest match, borderline case, and furthest match responses}
\label{borderline-response-distribution-e2}
\end{figure*}

@@

The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies. The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number. We used the same stimuli as in experiment 2. Table \ref{instructionse3} shows the instructions. % KvD I think we should motivate our choice of materials (especially "far ..er")

\begin{table*}[htbp]
\begin{center}
\caption{Experiment 3 Instructions assuming 6,15,24 dots as the Item, and showing fewer instead of more}
\label{instructionse3}
\begin{tabular}{llll}
\toprule
vagueness&instruction format&selection task&instruction\\
\midrule
crisp & numeric&matching & Choose a square with 6 dots \\ 
crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
vague & numeric&matching & Choose a square with about 10 dots \\ 
vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
\bottomrule
\end{tabular}
\end{center}
\end{table*}

@@

38 volunteers were recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds each for participating.
The apparatus was the same as Experiment 1.
The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{instructionse3}).
Each stimulus was an instruction followed by a triple of dots.
First a referring expression instruction was presented. Participants pressed a key to dismiss the instruction and proceed to the squares with dots in them.

@@

\begin{figure*}[htbp]
\centering
\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1.pdf}
\caption{Results for Experiment 3}
\label{resultse3}
\end{figure*}

\section{Experiment 4}
\subsection{Introduction}
This experiment investigated response times for instructions that did not use a number. 
We manipulated vagueness and the selection task (comparison and matching). 
In order to implement the experiment without mentioning numbers,  we used a prime to visually show the numbers of dots that we wanted to refer to as {\em the target} in the instructions. 
This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \textcite{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.
An item was thus a combination of a visual prime, a numeric triple, and a referring expression.
The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e4}.

\begin{table*}[htbp]
\centering
\caption{Experiment 4: Instructions}
\label{Instructions for e4}
\begin{tabular}{lllp{7cm}}
\hline
vagueness&instruction format& selection task&instruction\\
\hline
crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
crisp & verbal&comparison& Choose a square with fewer dots than the target \\
vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
vague & verbal&compaison& Choose a square with far fewer dots than the target \\ 
\hline
\end{tabular}
\end{table*}

@@

\begin{figure*}[htbp]
\centering
\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1-e4.pdf}
\caption{Results for Experiment 4}
\label{resultse4}
\end{figure*}

@@

\printbibliography

\appendix
\section{}

Table \ref{AllInstructions} shows the relevant properties of all the experiments in a single view.

\newgeometry{bottom=2cm}

\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
\newcommand\B{\rule[-2.2ex]{0pt}{0pt}} % Bottom strut

\begin{sidewaystable}
\caption{All Instructions}
\label{AllInstructions}

    \resizebox{\linewidth}{!}{%
    \begin{tabular}{clcccccc}
%    \midrule
    $E$&Example instruction&$V$&$Q$&Symmetry&Algorithm&$N$&Def\\
       \toprule
    1&Choose the square with 5 dots&Crisp&Number&Symmetric&Matching&1&Definite\\
    1&Choose the square with few dots&Vague&Word&Asymmetric&Comparison&1&Definite\B\\
%    \midrule
%\rule{0pt}{0cm} \\
    2&Choose the square with 6 dots&Crisp&Number&Symmetric&Matching&1&Definite\\
    2&Choose a square with about 10 dots&Vague&Number&Symmetric&Matching&2&Indefinite\\
    2&Choose the square with the fewest dots&Crisp&Word&Asymmetric&Comparison&1&Definite\\
    2&Choose a square with few dots&Vague&Word&Asymmetric&Comparison&2&Indefinite\B\\
%    \midrule
    3&Choose a square with 6 dots&Crisp&Number&Symmetric&Matching&1&Indefinite\\
    3&Choose a square with fewer than 20 dots&Crisp&Number&Asymmetric&Comparison&2&Indefinite\\
    3&Choose a square with about 10 dots&Vague&Number&Symmetric&Matching&2&Indefinite\\
    3&Choose a square with far fewer than 20 dots&Vague&Number&Asymmetric&Comparison&1&Indefinite\B\\
%    \midrule
    4&Choose a square with the same number of dots as the target&Crisp&Word&Symmetric&Matching&1&Indefinite\\
    4&Choose a square with fewer dots than the target&Crisp&Word&Asymmetric&Comparison&2&Indefinite\\
    4&Choose a square with about the same number of dots as the target&Vague&Word&Symmetric&Matching&2&Indefinite\\
    4&Choose a square with far fewer dots than the target&Vague&Word&Asymmetric&Comparison&1&Indefinite\\
    \bottomrule
    \end{tabular} 
    }
    ~\\
    E:~Experiment number\\
    $V$:~Vagueness\\
    $Q$:~Quantifier used\\
%    Symmetry:~\\
%    Algorithm:~\\
     $N$:~Number of valid targets for the instruction\\
     Def:~Definiteness of the article in the instruction: a/the

\end{sidewaystable}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{}

\end{document}

%\begin{figure*}
\includegraphics[angle=90,origin=c, scale=.5]{images/appendix_table}
%\end{figure*}
